---
title: "STA141 Assignment 2"
author: "RanLiu"
date: "October 12, 2015"
output: word_document
---
# Step One

1. function samedate() reads one txt file and returns a dataframe with date,lat,long and certain variable.
 
```{r, message=FALSE, queit = TRUE}
# function samedate() reads one txt file and returns a dataframe with four variables.
samedate = function(fullname){
  setwd("E:/RLiu_R/sta141")
  long = scan(fullname, skip = 5, nlines = 2, what = character(), quiet = TRUE)[1:24]
  date = scan(fullname, skip = 4, nlines = 1, what = character(), quiet = TRUE)[3]
  g = read.table(fullname, skip = 7, stringsAsFactors = FALSE)[,c(-2,-3)]
  colnames(g) = c("lat", long)
  library(reshape2)
  valueName = gsub("[0-9].*", "", basename(fullname))
  g = melt(g, id.vars = "lat", variable.name = "long", value.name = valueName)
  g$long = as.character(g$long)
  g$date = date
  g
}
```

```{r, queit = TRUE}
head(samedate("NASA/cloudhigh1.txt"))
```

2. combine different dates of same variable & do this for all the seven variables
+ get filenames, loop it over function samedate() to get all 72 data for one variable
+ loop over grouped filenames and get the similar data frame for all 7 variables.

```{r,  queit = TRUE}
#combine different dates of same variable & do this for all the seven variables
setwd("E:/RLiu_R/sta141")
varnames = gsub("[0-9].*", "", list.files("NASA/", pattern = "\\.txt$"))
groupfilenames = split(list.files("NASA/", pattern = "\\.txt$", full.names = TRUE), varnames)
samevar = function(txtnames){
  df = do.call(rbind, lapply(txtnames, samedate))
  return(df)
}
listofdfs = lapply(groupfilenames, samevar)
sapply(listofdfs, dim)
```

```{r, queit = TRUE}
head(listofdfs[[1]],3)
```

# Step Two

1. Check if locations and dates match across the seven variables.
+ function checktwo() checks whether lat long and date of two dataframes are identical
+ use checktwo() to compare each element of listofdfs with the first one
```{r,  queit = TRUE}
checktwo = function(x, y) {
  cols = c("lat", "long", "date")
  identical(x[, cols], y[, cols])
}
dfs_match = sapply(listofdfs[-1], checktwo, listofdfs[[1]])
all(dfs_match)
```

2.  Merge the seven dataframes into one big dataframe. Merge by same location and date.
+ function myMerge() merge two data frames by lat long and date
+ use Reduce() to loop over listofdfs and apply function myMerge()
```{r,  queit = TRUE}
myMerge = function (x, y){
  merge(x, y, all = TRUE, by = c("lat", "long", "date"))
}
whole = Reduce(myMerge, listofdfs)
```
```{r}
head(whole)
```

# Step Three: Elevation
1.  get elevations in each single grid, assign to "elev"
```{r,  queit = TRUE}
#get elevations in each single grid, assign to "elev"
setwd("E:/RLiu_R/sta141")
elev = read.table("NASA/intlvtn.dat", check.names = FALSE)
elev$lat = rownames(elev)
library(reshape2)
elev = melt(elev, id.vars = "lat", variable.name = "long", value.name = "elevation")
elev$long = as.character(elev$long)
```
```{r, queit = TRUE}
head(elev)
```

2. unify the latitude and longitude format
+ directions to +/-
+ round up/down lats and longs in "elev" to agree with those in "whole".
```{r,  queit = TRUE}
#unify the latitude and longitude format
##directions to +/-
###latitude
latchar = nchar(whole$lat)
directionNS = substr(whole$lat, latchar, latchar)
southern = (directionNS == "S")
whole$lat = substr(whole$lat, 1, latchar - 1)
whole$lat[southern] = - as.numeric(whole$lat[southern])
###longitude
longchar = nchar((whole$long))
directionEW = substr(whole$long, longchar, longchar)
western = (directionEW == "W")
whole$long = substr(whole$long, 1, longchar - 1)
whole$long[western] = - as.numeric(whole$long[western])
##round up/down lats and longs in "elev" to agree with those in "whole".
elev[, c("lat", "long")] = apply(elev[, c("lat", "long")], 2, as.numeric)
elev[, c("lat","long")] = round(elev[, c("lat","long")], 1)
```
```{r, queit = TRUE}
head(whole[, c("lat", "long")], 3)
```

3. match the lat and long and return corresponding elevation to each grid.
+ match() returns the indice of corresponding elevation in elev
+ use output mymatch to create a 41472 repeatance of elev with 72 rows.
```{r,  queit = TRUE}
##match the grid and add corresponding elevation to each row
whole$position = paste0(whole$lat, whole$long)
elev$position = paste0(elev$lat, elev$long)
mymatch = match(whole$position, elev$position)
whole$elev = elev$elevation[mymatch]
```
```{r, queit = TRUE}
head(whole, 3)
```
# Step Four: Exploring the Data Graphically

## Plot temperature versus pressure. Color code the points by the value of cloudlow.
1. break cloudlow into groups for coloring 
2. coerce the three variables are numeric
3. use ggplot() to touch modify it.
```{r, warning = FALSE, message = FALSE, fig.width = 12, fig.height = 9,  queit = TRUE}
library(ggplot2)
library(dplyr)
cloudlowGroup = cut(as.numeric(whole$cloudlow), breaks = 5)
whole[whole$cloudlow != "....", ] %>% 
  mutate(cloudlowGroup = cut(as.numeric(.$cloudlow), breaks = 5)) %>% 
  ggplot(aes(x = jitter(pressure), y = jitter(temperature), color = cloudlowGroup)) +
  geom_point(size = 2) +
  xlab("Pressure") +
  ylab("Temparature") +
  ggtitle("Temperature vs Pressure") +
  theme_bw() +
  theme(axis.title.y = element_text(size = 7))
```

## For the points at the four corners of the spatial grid, display the values of temperature over time.

1. find the lats and longs at the four corners
2. Coerce date to type Date
3.
```{r, fig.width = 12, fig.height = 9,  queit = TRUE}
#find the corners
whole[, c("lat", "long")] = apply(whole[, c("lat", "long")], 2, as.numeric)
bound =   function(x){
    min = min(x)
    max = max(x)
    return(c(min, max))
  }
bounds = as.data.frame(sapply(whole[,c("lat", "long")], bound))
corners = as.data.frame(expand.grid(bounds$lat, bounds$long))
colnames(corners) = c("lat", "long")
#date format
whole$date = as.Date(whole$date, "%d-%b-%Y")
#subset data of the four corners
cornerdata = whole[whole$lat %in% corners$lat & whole$long %in% corners$long, c("lat", "long", "temperature", "date")]
#plotting
cornerdata %>% 
  mutate(position = paste0("lat=", lat, " long=", long)) %>% 
  ggplot(aes(x = date, y = temperature, color = position)) +
  geom_line(size = 1) +
  facet_wrap(~position) +
  ggtitle('Temparater vs Date of Four Corners') +
  theme(legend.position="none") +
  theme_bw()
```

## For all points on the grid, compute the average and standard deviation for each of the 7 variables across time.
1. loop over each data frame's columns to calculate their mean.
2. loop over filenames :do the above for all seven variables.
3. merge the list result to get a data frame.
```{r, warning = FALSE, queit = TRUE}
#mean
whole$cloudlow = as.numeric(whole$cloudlow)
roundmean = function(x){round(mean(x), 2)}
myMean = function(x){
 aggregate(whole[,x] ~ whole$lat + whole$long, data = whole, roundmean)
}
wholemeans = lapply(unique(varnames), myMean)
merge2 = function(x, y){
  merge(x, y, all = TRUE, by = c("whole$lat", "whole$long"))
}
wholemeans = Reduce(merge2, wholemeans)
colnames(wholemeans) = c(c("lat", "long"), unique(varnames))
#sd
whole$cloudlow = as.numeric(whole$cloudlow)
mysd= function(x){
 aggregate(whole[,x] ~ whole$lat + whole$long, data = whole, sd)
}
wholesd = lapply(unique(varnames), mysd)
merge2 = function(x, y){
  merge(x, y, all = TRUE, by = c("whole$lat", "whole$long"))
}
wholesd = Reduce(merge2, wholesd)
colnames(wholesd) = c(c("lat", "long"), unique(varnames))
```

```{r, queit = TRUE}
head(wholemeans, 3)
head(wholesd, 3)
```

## Display the average value for pressure computed in the previous question on a map.
1. get a map using get_map, set parameters
2. add points to it
```{r, fig.width = 12, fig.height = 9,  warning = FALSE, queit = TRUE}
head(wholemeans)
library(ggmap)
mapbound = c(bounds$long[1], bounds$lat[1], bounds$long[2], bounds$lat[2])
hw2map = get_map(location = mapbound, source = "google", maptype = "terrain", color = "bw")
Pressure = wholemeans$pressure
ggmap(hw2map) +
  geom_point(aes(x = wholemeans$long, y = wholemeans$lat, color = Pressure), data = wholemeans, alpha = 1, size = 5) +
  theme_bw() +
  ggtitle("Average Pressure at Each Position")
```

## Display average surface temperature versus elevation.
*merge wholemeans data frame with elev, get meannelev.
```{r, fig.width = 12, fig.height = 9,  warning = FALSE, queit = TRUE}
meannelev = merge(elev, wholemeans, by = c("lat", "long"))
head(meannelev, 3)
ggplot(meannelev, aes(x = jitter(surftemp), y = jitter(elevation))) +
  geom_point() +
  geom_smooth() +
  xlab("Surface Temperature") +
  ylab("Elevation") +
  theme_bw() +
  ggtitle("Average Surface Temperature vs Elevation")
```

