---
title: "STA141 Assignment 5"
author: "RanLiu"
date: "November 21, 2015"
output: word_document
---

```{r}
# set up
setwd("E:/RLiu_R/sta141/HW5")
library(RSQLite)
db = dbConnect(SQLite(), "lean_imdbpy.db")
dbGetQuery(db, "SELECT * FROM genres " )
dbListTables(db)
dbListFields(conn = db, name = "genres")
```

1. How many actors are there in the database? How many movies?

For actors:
* Count personID in castInfo 
* join castInfo with roleType
* restrict role to only actor and actress.

```{r}
actor.total = dbGetQuery(db, "SELECT COUNT(DISTINCT(cast_info.person_id)) 
                         FROM cast_info, role_type
                         ON cast_info.role_id = role_type.id
                         WHERE role_type.id IN ('actor', 'actress)" )

```

For movies:
* count ID in title
* join title and kind_type
* restrict kind to only movie (as opposed to tv series or tv movie or other kinds).

```{r}
dbGetQuery(db, "SELECT * FROM title" )
movie.total = dbGetQuery(db, "SELECT COUNT(title.id)) 
                         FROM title, kindtype
                         WHERE title.kind_id = kind_type.id
                         AND kind_type.kind = 'movie'")
```

2. What time period does the database cover? 

Look through tables to find all fields that may contains date.

* From infoType table, we can see there are 6 types of info relating to date. These info are stored in the info field in the movieInfo table, movieInfoIDX table and personInfo table, all indexed by infoTypeID. They are: "release dates", "birth date", "death date", "LD date", "production dates", "filming dates"

```{r}
dbGetQuery(db, 'SELECT * FROM info_type
           WHERE info IN ("release dates", "birth date", "death date", 
                          "LD date", "production dates", "filming dates")')
```

Join the infoType table and 1.movieInfo table, 2.movieInfoIDX table and 3.personInfo table respectively.
And get the dates.

```{r}
dbGetQuery(db, 'SELECT movie_info.info, movie_info.info_type_id, info_type.info
           FROM movie_info, info_type
           WHERE movie_info.info_type_id = info_type.id
           AND info_type.info IN ("release dates", "birth date", "death date", 
                          "LD date", "production dates", "filming dates") LIMIT 10')
```


* From title table, we can see there is a "productionYear" field and a "seriesYears" field.

```{r}
dbGetQuery(db, 'SELECT MAX(production_year), series_years FROM title LIMIT 20')
```

3. What proportion of the actors are female? male?

* count id in name
* join name and roleType via castInfo
* restrict role to only actor and actress
* group by gender in name to get proportion

```{r}
dbGetQuery(db, 'SELECT COUNT(DISTINCT name.id)
           FROM name, cast_info, role_type 
           WHERE name.id = cast_info.person_id
           AND cast_info.role_id = role_type.id
           AND role_type.role IN ("actor", "actress")
           GROUP BY name.gender')
```

4. What proportion of the entries in the movies table are actual movies and what proportion are television series, etc.?

* count id/movieID ?????? in movieInfo
* join movieInfo and kindType via title
* group by kind in kindType

```{r}
dbGetQuery(db, "SELECT COUNT(DISTINCT(movie_id)) 
                         FROM movie_info, title, kindtype
                         WHERE movie_info.movie_id = title.id
                         AND title.kind_id = kind_type.id
                         GROUP BY kind_type.kind")
```

5. How many genres are there? What are their names/descriptions?

* count distinct genres in movieInfo
* join movieInfo and infoType
* retrict info in infoType to "genres"

```{r}
genres = dbGetQuery(db, "SELECT DISTINCT a.info
           FROM movie_info AS a, info_type AS b
           WHERE a.info_type_id = b.id
           AND b.info = 'genres' ")
#            info
# 1   Documentary
# 2    Reality-TV
# 3        Horror
# 4         Drama
# 5        Comedy
# 6       Musical
# 7     Talk-Show
# 8       Mystery
# 9          News
# 10        Sport
# 11       Sci-Fi
# 12      Romance
# 13       Family
# 14        Short
# 15    Biography
# 16        Music
# 17    Game-Show
# 18    Adventure
# 19        Crime
# 20          War
# 21      Fantasy
# 22     Thriller
# 23    Animation
# 24       Action
# 25      History
# 26        Adult
# 27      Western
# 28    Lifestyle
# 29    Film-Noir
# 30 Experimental
# 31   Commercial
# 32      Erotica
```

So there are 32 genres in total. And their names are listed above.

6. List the 10 most common genres of movies, showing the number of movies in each of these genres.

* create a new table by select from movieinfo that has genre info 
* count movieID
* group by info, which are all kinds of genres 

```{r}
dbGetQuery(db, "CREATE TABLE genres AS 
           SELECT a.info, a.movie_id
           FROM movie_info AS a, info_type AS b
           WHERE a.info_type_id = b.id
           AND b.info = 'genres' ")

dbGetQuery(db, "SELECT info, COUNT(movie_id)
           FROM genres
           GROUP BY info
           ORDER BY COUNT(movie_id) DESC
           LIMIT 10")
```

7. Find all movies with the keyword 'space'. How many are there? What are the years these were released? and who were the top 5 actors in each of these movies?

7(1). How many are there

* count movieID in movieKeyword
* join movieKeyword and keyword
* restrict keyword field in keyword table to "space"

```{r}
dbGetQuery(db, "SELECT COUNT(DISTINCT a.movie_id), b.keyword
           FROM movie_keyword AS a, keyword AS b
           WHERE a.keyword_id = b.id
           AND b.keyword = 'space' ")

## keyword COUNT(movie_id)
## space             932
```

There are 932 movies with keyword 'space'.

7(2). What are the years these were released?

* get release dates from movie_info (need processing to get only the years)
* join movieInfo and infoType to know which are info of "release dates" 
* join movieInfo and keyword via title and movie_keyword
* restrict keyword in keyword to 'space'

```{r}
dbGetQuery(db, "SELECT * FROM info_type
           WHERE info_type.info LIKE 'r%' ")
# connect via title 
dbGetQuery(db, "SELECT c.title, b.info, a.info, a.movie_id, e.keyword
           FROM movie_info AS a, info_type AS b, title AS c, 
           movie_keyword AS d, keyword AS e
           WHERE a.info_type_id = b.id
           AND b.info = 'release dates'
           AND a.movie_id = c.id
           AND c.id = d.movie_id
           AND d.keyword_id = e.id
           AND e.keyword = 'space' LIMIT 10")
```

7(3). who were the top 5 actors in each of these movies

* create a table to store the movieID of movies with keyword "space"
* join space table and name via castInfo
* find top actors group by movieID

```{r, "create table space"}
#create new table
dbGetQuery(db, "CREATE TABLE space AS 
           SELECT DISTINCT a.movie_id, b.keyword
           FROM movie_keyword AS a, keyword AS b
           WHERE a.keyword_id = b.id
           AND b.keyword = 'space' ")
# Have a look at the newly created table "space"
dbGetQuery(db, "SELECT * FROM space LIMIT 5")
```

```{r, "find top five actors"}
top.actor.space = dbGetQuery(db, "SELECT s.keyword, s.movie_id, c.nr_order, n.name
           FROM space AS s, cast_info AS c, name AS n
           WHERE s.movie_id = c.movie_id
           AND c.person_id = n.id
           AND c.nr_order < 6
           ORDER BY s.movie_id, c.nr_order")
```

8. Has the number of movies in each genre changed over time? Plot the overall number of movies in each year over time, and for each genre.

*       1800 1801 .... 2016
genre1  23    33 ....   456
genre2  23    33 ....   456
 ...
 
 title
 
```{r}

```















